---
Author: "Betsiet Delelegn"
title: "week-9-R"
---

## In-class Activity Week 9

#### Required Packages

```{r}
library(DBI)
library(dbplyr)
library(tidyverse)
```

#### Activity 2

```{r}
requireNamespace("RPostgreSQL")
db_conn <- dbConnect(dbDriver('PostgreSQL'),
                     host = "dkepplin-stat463.cec.gmu.edu",
                     dbname = "stat507cherry",
                     user = "stat507learner",
                     password = "learningSQLinSTAT507!")
str(db_conn)

dbListObjects(db_conn)
dbListFields(db_conn, "cherry_locations")
dbListFields(db_conn, "bloom_dates")
dbListFields(db_conn, "weather_stations")
dbListFields(db_conn, "weather_data")

```

The tables cherry_location, bloom_dates and weather_stations have location_id that can link them all together. The weather_data can be connected to weather_stations using noaa_station_id.

#### Activity 3

```{r}
cherry_locations_db <- tbl(db_conn, "cherry_locations")
bloom_dates_db <- tbl(db_conn, "bloom_dates")

cherry_locations_db |> 
  filter(country == "USA")

cherry_locations_db |> 
  mutate(country = if_else(country == "USA", "U.S.A.", country)) |> 
  filter(country == "U.S.A.") |>
  show_query()
```

The query gets all column from cherry-location and within the sub-query it replaces USA with U.S.A but leaves other countries unchanged.

#### Activity 4

```{r}
bloom_dates_db <- tbl(db_conn, "bloom_dates")
weather_stations_db <- tbl(db_conn, "weather_stations")

bloom_dates_db|>
  inner_join(weather_stations_db, by = "location_id")|>
  show_query()
```

Here the query is fully processed on the database server.

#### Activity 5

```{r}
weather_data_db <- tbl(db_conn, "weather_data")

bloom_dates_db |>
  inner_join(weather_stations_db, by = "location_id") |>
  inner_join(weather_data_db, by = "noaa_station_id") |>
  group_by(location_id, year.x) |>
  summarise(
    min(cldd), mean(cldd), max(cldd),
    min(dp01), mean(dp01), max(dp01),
    min(dt32), mean(dt32), max(dt32),
    min(dx32), mean(dx32), max(dx32),
    min(dx70), mean(dx70), max(dx70),
    min(htdd), mean(htdd), max(htdd),
    min(prcp), mean(prcp), max(prcp),
    min(tavg), mean(tavg), max(tavg),
    min(tmax), mean(tmax), max(tmax),
    min(tmin), mean(tmin), max(tmin),
  
  ) |> 
  show_query()
```
